qcircuit 2.5.3 Tutorial

### 2017-09-22 CO681-QuantumInformationProcessing/assignments/assignment_01.tex
problems 1 -3

\begin{document}
\section{}
\paragraph{}
Let $X = \begin{bmatrix} 0 & 1 \\ 1 & 0 \end{bmatrix}$, $H = \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 1 \\ 1 & -1 \end{bmatrix}$, $T = \begin{bmatrix} 1 & 0 \\ 0 & \zeta_8\end{bmatrix}$
\begin{itemize}
\item a) $\begin{bmatrix} 0 & 1 \\ \zeta_8 & 0 \end{bmatrix} = TX$
\item b) $\frac{1}{\sqrt{2}}\begin{bmatrix} 1 & 1 \\ -\zeta_8 & \zeta_8 \end{bmatrix} = THX$
\item c) $\frac{1}{\sqrt{2}}\begin{bmatrix} 1 &\zeta_8 \\ \zeta_8 & -i\end{bmatrix}=THT$
\item d) $\frac{1}{2} \begin{bmatrix} 1+\zeta_8 & 1-\zeta_8 \\ 1-\zeta_8 & 1+\zeta_8\end{bmatrix}=HTH.$
\end{itemize}
\section{}
\subsection{a}
\paragraph{}
Let $U$ be the unitary matrix for CNOT given by
$$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \\ 0& 0 & 1 &0 \end{bmatrix}.$$
The eigenvectors for $U$ are $\ket{00}$, $\ket{01}$, and $\ket{1+}$ all with eigenvalue $1$ and $\ket{1-}$ with eigenvalue $-1$. We have four orthonormal eigenvectors, so we've found all of them.
\subsection{b}
\paragraph{}
The circuit for $(H \otimes I)U$ is
\[ \Qcircuit @C=1em @R=1em {
 & \ctrl{1} & \gate{H} &  \qw \\
 & \targ &  \qw& \qw
}\]
and the corresponding unitary is
$$B:= \frac{1}{\sqrt{2}}\begin{bmatrix} 1 & 0 & 1 & 0 \\ 0 & 1 & 0 & 1 \\ 1 & 0 & -1 & 0 \\ 0 & 1 & 0 & -1 \end{bmatrix}\begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \\ 0& 0 & 1 &0 \end{bmatrix} = \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 0 & 0 & 1 \\ 0 & 1 & 1 & 0 \\ 1 & 0 & 0 & -1\\ 0 & 1 & -1 & 0\end{bmatrix}.$$
\subsection{c}
\paragraph{}
Let $W$ be the unitary matrix for NOTC given by
$$W = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 0 & 0 & 1 \\ 0 & 0 & 1 & 0 \\ 0& 1 & 0 &0 \end{bmatrix}.$$
The circuit for $(I \otimes H)W$ is
\[ \Qcircuit @C=1em @R=1em {
 & \targ & \qw &  \qw \\
 & \ctrl{-1} & \gate{H} & \qw
}\]
and the corresponding unitary is
$$C:= \frac{1}{\sqrt{2}}\begin{bmatrix} 1 & 1 & 0 & 0 \\ 1& -1 & 0 & 0 \\ 0 & 0 & 1 & 1 \\ 0& 0 & 1 &-1 \end{bmatrix}\begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 0 & 0 & 1 \\ 0 & 0 & 1 & 0 \\ 0& 1 & 0 &0 \end{bmatrix} = \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 0 & 0 & 1 \\ 1 & 0 & 0 & -1 \\ 0 & 1 & 1 & 0\\ 0 & -1 & 1 & 0\end{bmatrix}.$$
\subsection{d}
\paragraph{}
Since $(H\otimes H)U(H\otimes H) = W$ if and only if 
$$(I \otimes H)(H \otimes H) U(H \otimes H) = (I\otimes H) W$$
which simplifies as
$$(H\otimes I) U(H\otimes H) = (I \otimes H) W,$$
it suffices to show that $B (H\otimes H) = C$. Thus we compute
\begin{align*}B(H\otimes H) &= \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 0 & 0 & 1 \\ 0 & 1 & 1 & 0 \\ 1 & 0 & 0 & -1\\ 0 & 1 & -1 & 0\end{bmatrix}\cdot \frac{1}{2}\begin{bmatrix} 1 & 1 & 1 & 1 \\ 1 & - 1 & 1 & -1 \\ 1 & 1& -1 & -1 \\ 1 & -1 & -1 &1\end{bmatrix} \\
&=\frac{1}{\sqrt{2}}\cdot \frac{1}{2}\begin{bmatrix}2 &0 & 0 & 2 \\ 2 & 0 & 0 & -2 \\ 0 & 2 & 2 & 0 \\ 0 & -2 & 2 & 0 \end{bmatrix} \\
& C. \blacksquare\end{align*}
\end{document}

>>>>

\begin{document}
\section{}
\paragraph{}
Let $X = \begin{bmatrix} 0 & 1 \\ 1 & 0 \end{bmatrix}$, $H = \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 1 \\ 1 & -1 \end{bmatrix}$, $T = \begin{bmatrix} 1 & 0 \\ 0 & \zeta_8\end{bmatrix}$
\begin{itemize}
\item a) $\begin{bmatrix} 0 & 1 \\ \zeta_8 & 0 \end{bmatrix} = TX$
\item b) $\frac{1}{\sqrt{2}}\begin{bmatrix} 1 & 1 \\ -\zeta_8 & \zeta_8 \end{bmatrix} = THX$
\item c) $\frac{1}{\sqrt{2}}\begin{bmatrix} 1 &\zeta_8 \\ \zeta_8 & -i\end{bmatrix}=THT$
\item d) $\frac{1}{2} \begin{bmatrix} 1+\zeta_8 & 1-\zeta_8 \\ 1-\zeta_8 & 1+\zeta_8\end{bmatrix}=HTH.$
\end{itemize}
\section{}
\subsection{a}
\paragraph{}
Let $U$ be the unitary matrix for CNOT given by
$$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \\ 0& 0 & 1 &0 \end{bmatrix}.$$
The eigenvectors for $U$ are $\ket{00}$, $\ket{01}$, and $\ket{1+}$ all with eigenvalue $1$ and $\ket{1-}$ with eigenvalue $-1$. We have four orthonormal eigenvectors, so we've found all of them.
\subsection{b}
\paragraph{}
The circuit for $(H \otimes I)U$ is
\[ \Qcircuit @C=1em @R=1em {
 & \ctrl{1} & \gate{H} &  \qw \\
 & \targ &  \qw& \qw
}\]
and the corresponding unitary is
$$B:= \frac{1}{\sqrt{2}}\begin{bmatrix} 1 & 0 & 1 & 0 \\ 0 & 1 & 0 & 1 \\ 1 & 0 & -1 & 0 \\ 0 & 1 & 0 & -1 \end{bmatrix}\begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \\ 0& 0 & 1 &0 \end{bmatrix} = \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 0 & 0 & 1 \\ 0 & 1 & 1 & 0 \\ 1 & 0 & 0 & -1\\ 0 & 1 & -1 & 0\end{bmatrix}.$$
\subsection{c}
\paragraph{}
Let $W$ be the unitary matrix for NOTC given by
$$W = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 0 & 0 & 1 \\ 0 & 0 & 1 & 0 \\ 0& 1 & 0 &0 \end{bmatrix}.$$
The circuit for $(I \otimes H)W$ is
\[ \Qcircuit @C=1em @R=1em {
 & \targ & \qw &  \qw \\
 & \ctrl{-1} & \gate{H} & \qw
}\]
and the corresponding unitary is
$$C:= \frac{1}{\sqrt{2}}\begin{bmatrix} 1 & 1 & 0 & 0 \\ 1& -1 & 0 & 0 \\ 0 & 0 & 1 & 1 \\ 0& 0 & 1 &-1 \end{bmatrix}\begin{bmatrix} 1 & 0 & 0 & 0 \\ 0& 0 & 0 & 1 \\ 0 & 0 & 1 & 0 \\ 0& 1 & 0 &0 \end{bmatrix} = \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 0 & 0 & 1 \\ 1 & 0 & 0 & -1 \\ 0 & 1 & 1 & 0\\ 0 & -1 & 1 & 0\end{bmatrix}.$$
\subsection{d}
\paragraph{}
Since $(H\otimes H)U(H\otimes H) = W$ if and only if 
$$(I \otimes H)(H \otimes H) U(H \otimes H) = (I\otimes H) W$$
which simplifies as
$$(H\otimes I) U(H\otimes H) = (I \otimes H) W,$$
it suffices to show that $B (H\otimes H) = C$. Thus we compute
\begin{align*}B(H\otimes H) &= \frac{1}{\sqrt{2}} \begin{bmatrix} 1 & 0 & 0 & 1 \\ 0 & 1 & 1 & 0 \\ 1 & 0 & 0 & -1\\ 0 & 1 & -1 & 0\end{bmatrix}\cdot \frac{1}{2}\begin{bmatrix} 1 & 1 & 1 & 1 \\ 1 & - 1 & 1 & -1 \\ 1 & 1& -1 & -1 \\ 1 & -1 & -1 &1\end{bmatrix} \\
&=\frac{1}{\sqrt{2}}\cdot \frac{1}{2}\begin{bmatrix}2 &0 & 0 & 2 \\ 2 & 0 & 0 & -2 \\ 0 & 2 & 2 & 0 \\ 0 & -2 & 2 & 0 \end{bmatrix} \\
& C. \blacksquare\end{align*}
\section{}
\paragraph{}
Let $\ket{\psi} = \frac{1}{\sqrt{2}}\ket{00} + \frac{1}{2}\ket{01} + \frac{1}{2}\ket{11}.$
\subsection{a}
\paragraph{}
When we measure the first qubit with respect to the computational basis ($\ket{0}, \ket{1}$) we obtain $\ket{0}$ with probability
$$|\frac{1}{\sqrt{2}}|^2 + |\frac{1}{2}|^2 = \frac{3}{4}$$
and the second qubit would be left in the normalized post-measurement state
$$\sqrt{\frac{2}{3}} \ket{0} + \frac{1}{\sqrt{3}} \ket{1},$$
and we obtain $\ket{1}$ with probability
$$|\frac{1}{2}|^2 = \frac{1}{4}$$
and the second qubit would be left in the normalized post-measurement state
$$\ket{1}.$$
\subsection{b}
\paragraph{}
Rewriting $\ket{\psi}$ in the $\ket{+}, \ket{-}$ basis gives
$$\ket{\psi} = \frac{\sqrt{2} + 2}{4}\ket{++} + \frac{\sqrt{2}-2}{4}\ket{+-} + \frac{\sqrt{2}}{4}\ket{-+} + \frac{\sqrt{2}}{4}\ket{--}.$$
So measuring the first qubit with respect to the $\ket{+}, \ket{-}$ basis gives $\ket{+}$ with probability
$$|\frac{\sqrt{2}+2}{4}|^2+ |\frac{\sqrt{2}-2}{4}|^2 = \frac{3}{4}.$$
and the second qubit would be left in the normalized post-measurement state
$$ \frac{\sqrt{2} + 2}{2\sqrt{3}}\ket{+} + \frac{\sqrt{2}-2}{2\sqrt{3}}\ket{-},$$
and gives $\ket{-}$ with probability
$$2|\frac{\sqrt{2}}{4}|^2 = \frac{1}{4}$$
and the second qubit would be left in the normalized post-measurement state
$$\frac{1}{\sqrt{2}}\ket{+} + \frac{1}{\sqrt{2}}\ket{-} .$$
\end{document}
